Exacerbated by the Pandemic: Analyzing the Impact of COVID-19 on Mortality Among Toronto's Homeless with a Lens on Gender Disparities

### 2024-03-16 paper/paper (2).qmd
Add files via upload

The analysis will be conducted in R (R Core Team 2020) and will utilize an array of packages such as tidyverse (Wickham et al. 2019), readr (Wickham, Hester, and Bryan 2024), dplyr (Wickham et al. 2021), and opendatatoronto (Gelfand 2020) for data manipulation and analysis. All visualizations in this report are generated through ggplot2 (Wickham 2016), with tables crafted using kableExtra (Zhu 2021).

@@

@@

Where:

- $Y_{it}$ is the death count for gender $i$ at time $t$,
- $T_t$ is a binary indicator for the time period, with $T = 0$ for the pre-COVID-19 period and $T = 1$ for the post-COVID-19 period,
- $G_i$ is a binary indicator for gender, with $G = 1$ for one gender (e.g., females) and $G = 0$ for the other (e.g., males),
- $\delta$ represents the DiD estimator, capturing the interaction effect between the time period and gender, indicating the differential impact of COVID-19 on death counts across genders,
- $\mathbf{X}_{it}$ is a vector of control variables for gender $i$ at time $t$,
- $\boldsymbol{\theta}$ is a vector of coefficients for the control variables,
- $\alpha$, $\beta$, and $\gamma$ are coefficients,
- $\epsilon_{it}$ is the error term.

>>>>

The analysis will be conducted in [@citeR] and will utilize an array of packages such as [@citetidyverse], [@ciitereadr], [@citedplyr], and [@citeopendatatoronto] for data manipulation and analysis. All visualizations in this report are generated through [@citeggplot2], with tables crafted using [@citekableExtra].

@@

## Model Setup

@@

## Model Justification

- $alpha$ is constant intercept term of the model, capturing the dependent variable(death counts) when all independent variables are 0.
- $Y_{it}$ represents the dependent variable, which in this context is the number of deaths among the homeless population for individual $i$ at time $t$, 
- $T_t$ is a binary indicator for the time period, with $T = 0$ for the pre-COVID-19 period and $T = 1$ for the post-COVID-19 period,
- $G_i$ is a binary indicator for gender, with $G = 1$ for one gender (e.g., females) and $G = 0$ for the other (e.g., males),
- $\delta$ represents the DiD estimator, capturing the interaction effect between the time period and gender, indicating the differential impact of COVID-19 on death counts across genders,
- $\mathbf{X}_{it}$ is a vector of control variables for gender $i$ at time $t$,
- $\boldsymbol{\theta}$ is a vector of coefficients for the control variables,
- $\alpha$, $\beta$, and $\gamma$ are coefficients of time, gender and interaction of time and gender.
- $\epsilon_{it}$ is the error term, capturing the unobserved influences on the dependent variable.

The inclusion of the interaction term $(T_t \times G_i)$ is particularly important as it acknowledges that the relationship between time (and therefore the impact of COVID-19) and death counts may not be uniform across genders. This model is justifiable as it provides a comprehensive framework for examining the potential changes in homeless death counts over time while considering the interaction between time and gender, as well as the effect of other relevant factors. 
This model allows for a deeper understanding of the dynamics at play in this critical public health issue.
